---
title: "Practical Machine Learning CP"
output: html_document
---

Presented here is an analysis of a data set generated from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.  The goal of this analysis is to build a model/classifier that will take a set of observations as input and predict the manner (classe) in which the exercise was performed for each observation.

***Loading the dataset***
```{r}
curls<-read.table("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", header=T, sep=",", na.strings=c("NA", "", "#DIV/0!"))
```

***Cleaning the dataset***

```{r}
#Remove any variable that contains NAs
nacount<-sapply(curls, is.na)
nacount<-as.data.frame(nacount)
totals<-sapply(nacount, sum)
ColumnsToRemove<-which(totals>0)
curls<-curls[,-ColumnsToRemove]
#Remove any variable related to time because this isn't likely to affect classe
curls<-curls[, -grep("time", names(curls))]
#Manual inspection of remaining variables suggests that the first 4 variables can also be removed
curls<-curls[,-(1:4)]
```

***Partition training data into a training and testing set***
```{r}
library(caret)
partition<-createDataPartition(curls$classe, p=.7, list=F)
curltrain<-curls[partition,]
curltest<-curls[-partition,]
```

***A Random Forest Model***
```{r}
#We are trying to predict a categorical variable that has multiple levels.  In many cases like this Random Forests are a good approach.  Therefore, the following code trains a random forest model using 50 trees and 3-fold cross validation.  The choice of 50 trees and 3-fold cross validation reduces computing time while still generating an accurate model

modFit<-train(classe~., data=curltrain, method="rf", ntree=50, number=3)
#Let's take a look at the model
modFit
#We can see that the final model chosen has an accuracy of 98.7% on the training data.  This appears to be a very strong model.  Let's see how it performs on the test data and get an estimation of the out of sample error rate
predictions<-predict(modFit, newdata=curltest)
confusionMatrix(predictions, curltest$classe)
#We can see the accuracy on the test set is very high (>99%).  This provides further evidence for the strength of this model and suggests that the out of sample error rate will be very low (<1%).  Further support for the validity of this model as a strong classifier comes from the 20/20 score that it acheived on the 20 test samples provided for the programming portion of this assigment.
```