---
title: "Practical Machine Learning CP"
output: html_document
---

Presented here is an analysis of a data set generated from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.  The goal of this analysis is to build a model/classifier that will take a set of observations as input and predict the manner (classe) in which the exercise was performed for each observation.

***Loading the dataset***
```{r}
curls<-read.table("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", header=T, sep=",", na.strings=c("NA", "", "#DIV/0!"))
```

***Cleaning the dataset***

```{r}
#Remove any variable that contains NAs
nacount<-sapply(curls, is.na)
nacount<-as.data.frame(nacount)
totals<-sapply(nacount, sum)
ColumnsToRemove<-which(totals>0)
curls<-curls[,-ColumnsToRemove]
#Remove any variable related to time data was recorded as this won't affect classe
curls<-curls[, -grep("time", names(curls))]
#Manual inspection of remaining variables suggests that the first 4 variables can also be removed
curls<-curls[,-(1:4)]
```

***Partition training data into a training and testing set***
```{r}
library(caret)
partition<-createDataPartition(curls$classe, p=.7, list=F)
curltrain<-curls[partition,]
curltest<-curls[-partition,]
```

***A Random Forest Model***
```{r}
#We are trying to predict a categorical variable that has multiple levels.  #Random forests are typically a good approach in such cases.  The following
#code trains a random forest model using 50 trees and 3-fold cross
#validation.  The choice of 50 trees and 3-fold cross validation reduces 
#computing time while still generating an accurate model

modFit<-train(classe~., data=curltrain, method="rf", ntree=50, number=3)
modFit
modFit$finalModel
predictions<-predict(modFit, newdata=curltrain)
confusionMatrix(predictions, curltrain$classe)
predictions<-predict(modFit, newdata=curltest)
confusionMatrix(predictions, curltest$classe)
```